I thank \textbf{Dr. Deryle Lonsdale} for his comprehension of computational linguistics and his ability to teach it with zeal and effectiveness. He kept me in my seat, awestruck by how amazing studying syntax could be.\bigskip \medskip

I thank \textbf{Dr. Eric Ringger} for his comprehension of natural language processing and his ability to teach critically rather than by rote. He tackled every question with ease and foresight.\bigskip \medskip

I thank \textbf{Dr. Alan K. Melby} for his vigor and zeal. He taught me that linguistics can be fun, even if computers are not involved, teaching me that even the humanities side of things can leave you thinking for days.\bigskip \medskip %Who would have thought?

I thank \textbf{Dr. Mark Davies} for being a friend and colleague. Being able to participate in helping prepare the corpora for GloWbE \citep{glowbe}, one of the largest English dialect corpora in the world, was an amazing experience to say the least!\bigskip \medskip %-Extracting and learning from large amounts of data is what he does.-

%I thank LoriAnne Spear for being a great department secretary, always keeping things in order, ...etc..\bigskip \medskip % TODO: Write more on LoriAnne.
%\\

I wish to thank my wife, \textbf{Wendy Bean}, who helped me to choose a field of study and supported me along the way. %as well as survive in it.
She has been my friend, peer, and confidant. She has proven to be true help-meet for me, especially when times get tough (such as when a database gets corrupted or locks up or when I think you are a victim of an amazingly pernicious virus and is willing to help you change all your passwords at a moment's notice.) She's also good at forgetting the passwords, which makes them even more secure...\bigskip \medskip

I thank my \textbf{daughter} for spilling cereal on me when I needed a reason to take a break, and for being one of my best friends.% Additionally, for teaching me that linguistics is more fun than you think.Apparently, ``am:a:m!'' means ``Mom''---or any parental figure for that matter.\bigskip \medskip

Last, but by far not least, I also thank \textbf{God} for his help.\bigskip \medskip

% TODO: Or just simply use:
%I thank all the professors and teachers I have had along the way, even the supervisors I have at work. I especially thank my wife who knew how to support me in the process whether it be file corruption recovery or otherwise.
